---
title: "First Rmarkdown"
author: "Mel Schwan"
date: "5/13/2019"
output: html_document
---
## Median Histogram from Random Exponentials

# Introduction 
Code to loop through a calculation of the medians of random vector being generated.
 n is the number of exponential samples taken and nsim is how many times we loop through 
 our calculations of medians.

# Specifications
  * Create lotsa.medians variable
  * Loop through nsim times
  * Create exponential random generation of numbers and assign to x
  * Calculate the median of array.
  * Assign calculations to variables for later calculations.
  * Plot histogram of exponential medians calculation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#R Code for Unit 1 by Mel Schwan 05/12/2019
```{r}
n <- 5
nsim <- 1000

lotsa.medians <- numeric(nsim)



for (i in 1:nsim) {
  x <- rexp(n)
  lotsa.medians[i] <- median(x)
}


sum_med = summary(lotsa.medians)
print(sum_med)
len_lotsa = length(lotsa.medians)
print(len_lotsa)
mean_lotsa = mean(lotsa.medians)
print(mean_lotsa)

## Including Plots

hist(lotsa.medians)
echo = TRUE
```
## Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
